---
title: 'Random Fractal Maps: Software Comparison'
author: ~
date: '2018-02-18'
slug: random-fractal-maps-software-comparison
categories: []
tags: []
output: md_document
---

Here I show how to use three different software packages to produce neutral landscape models via midpoint displacement. First, I show use of the newly released `NLMR` package which runs natively in `R`. Next, I show use of `GRASS` (via `r.surf.fractal`) and `nlmpy` using their respective `R` wrappers.

```{r message=FALSE}
library(NLMR)
library(raster)
library(boot)
library(rgrass7)
library(reticulate)
library(ggplot2)
library(purrr)
library(dplyr)
library(cowplot)
```

## NLMR package [https://github.com/marcosci/NLMR](https://github.com/marcosci/NLMR)

> The roughness value controls the level of spatial autocorrelation. 

```{r nlmr, message=FALSE}
nlmr_mpd <- function(ncol = 17, nrow = 17, roughness = 0.6) {
  nlm_mpd(ncol = ncol, nrow = nrow, roughness = roughness)
}
```

## GRASS r.surf.fractal [https://grass.osgeo.org/grass74/manuals/r.surf.fractal.html](https://grass.osgeo.org/grass74/manuals/r.surf.fractal.html)

> The dimension value controls the fractal dimension of surface (2 < D < 3).

```{r grass, message=FALSE, results='hide'}
initGRASS("/usr/lib/grass74", home = tempdir(), override = TRUE)

raster_template <- as(nlmr_mpd(), "SpatialGridDataFrame")

grass_mpd <- function(dimension = 2.05){
  writeRAST(raster_template, "grass_mpd", flags = "overwrite")
  execGRASS("g.region", raster = "grass_mpd")
  execGRASS("r.surf.fractal", output = "grass_mpd", flags = "overwrite", 
            dimension = dimension)
  res <- raster(readRAST("grass_mpd"))
  res <- raster::scale(res)
  res[] <- boot::inv.logit(res[])
  res
}
```

## NLMpy [https://pypi.python.org/pypi/nlmpy](https://pypi.python.org/pypi/nlmpy)

> The h value controls the level of spatial autocorrelation in element values.

```{r echo=FALSE}
use_condaenv("blog")
```

```{r }
nlmpy <- import("nlmpy")
nlmpy_mpd <- function(nRow = 18L, nCol = 18L, h = 0.6){
  res <- nlmpy$nlmpy$mpd(nRow = nRow, nCol = nCol, h = h)
  raster(res, ymx = nrow(res), xmx = ncol(res))
}
```


```{r comparison, message=FALSE, results='hide'}
# NLMR
nlmr_mpd_res <- lapply(1:5, function(x) 
                  as.data.frame(as(nlmr_mpd(), "SpatialPixelsDataFrame")))
names(nlmr_mpd_res) <- 1:5
nlmr_mpd_res <- map2_df(nlmr_mpd_res, names(nlmr_mpd_res), ~ mutate(.x, ID = .y))

# GRASS
grass_mpd_res <- lapply(1:5, function(x) 
                  as.data.frame(as(grass_mpd(dimension = 2.5),
                                   "SpatialPixelsDataFrame")))
names(grass_mpd_res) <- 1:5
grass_mpd_res <- map2_df(grass_mpd_res, names(grass_mpd_res), ~ mutate(.x, ID = .y))


# NLMpy
nlmpy_mpd_res <- lapply(1:5, function(x) 
                  as.data.frame(as(nlmpy_mpd(h = 0.8), "SpatialPixelsDataFrame")))
names(nlmpy_mpd_res) <- 1:5
nlmpy_mpd_res <- map2_df(nlmpy_mpd_res, names(nlmpy_mpd_res), ~ mutate(.x, ID = .y))

# Viz
plot_grid(
  ggplot() + 
    geom_tile(data = nlmr_mpd_res, aes(x = x, y = y, fill = layer)) + 
    scale_fill_viridis_c() + 
    facet_wrap(~ID, nrow = 1) + theme_nothing(),
  ggplot() + 
    geom_tile(data = grass_mpd_res, aes(x = x, y = y, fill = grass_mpd)) + 
    scale_fill_viridis_c() + 
    facet_wrap(~ID, nrow = 1) + theme_nothing(),
  ggplot() + 
    geom_tile(data = nlmpy_mpd_res, aes(x = x, y = y, fill = layer)) + 
    scale_fill_viridis_c() + 
    facet_wrap(~ID, nrow = 1) + theme_nothing(), 
nrow = 3, ncol = 1, labels = c("NLMR", "GRASS", "NLMpy"), hjust = -0.1, vjust = 2)
```

I played with the `dimension` argument of `grass_mpd()` and the `h` argument of `nlmpy_mpd()` to try to get a result that was qualitatively similar to the results of `nlmr_mpd()`. Note that I have set the size the the `nlmpy` raster slightly smaller. Setting it to `17L` causes a strange error that I am hesitant to dig into. 

Each of the three options analyzed here have a different specification for controlling  surface _roughness_. This makes analytical comparisons difficult. `NLMR` definitely is the easiest to use but I suspect that its native `R` implementation would make it much slower than `GRASS` for large rasters. `NLMpy` is also easy to use but the errors I encountered when running a 17 x 17 matrix makes me hesitant to recommend it.

```{shell, eval=FALSE, echo=FALSE}
# setup
pip install nlmpy
```