---
title: Automated roxygen documentation of R package data
author: null
date: '2020-07-12'
slug: automated-roxygen-documentation-of-r-package-data
categories: []
tags: []
editor_options: 
  chunk_output_type: console
---

Documenting data is an important part of the R package publishing process. Most packages document their data using a named list so that they have a roxygen block that looks like this:

```
#' \describe{
#'   \item{One}{Description of the One variable}
#'   \item{Two}{Description of the Two variable}
#' }
```

Writing out named lists in this way is fine for small and infrequently updated data. However, the manual entry and non-tabular formatting makes for difficult reuse in other contexts such as in [data dictionaries](https://www.usgs.gov/products/data-and-tools/data-management/data-dictionaries). We can make the storage and entry of data documentation cleaner and enable reuse by storing the data dictionary in a separate file from the R source code. A relatively simple way to do this is to pass your data dictionary through a function to convert it into a LaTeX-style `\tabular` object that can be manually pasted into a roxygen block (See http://r-pkgs.had.co.nz/man.html):

```{r }
# modified from http://r-pkgs.had.co.nz/man.html
tabular <- function(df, ...) {
  stopifnot(is.data.frame(df))

  align <- function(x) if (is.numeric(x)) "r" else "l"
  col_align <- vapply(df, align, character(1))

  cols <- lapply(df, format, ...)
  contents <- do.call("paste",
    c(cols, list(sep = " \\tab ", collapse = "\\cr\n  ")))

  paste("\\tabular{", paste(col_align, collapse = ""), "}{\n  ",
    contents, "\n}\n", sep = "")
}
```

The output would be manually copy-pasted into roxygen code looking like this:

```{r echo=FALSE}
dictionary <- dplyr::bind_rows(
  c(name = "One", description = "Description of the One variable"),
  c(name = "Two", description = "Description of the Two variable")
)
```

```{r }
cat(tabular(dictionary))
```

The problem with this approach is that the interactive manual copying becomes extremely cumbersome if you have a lot of different datasets to document. I discovered an automated solution using (a little known) `@eval` roxygen tag. What follows is a a step-by-step guide to creating automated roxygen data documentation that pulls from an external data dictionary. For this demonstration we will use the [`population` dataset](https://tidyr.tidyverse.org/reference/who.html) contained in the `tidyr` R package. This strategy is implemented in a fully functioning R package at https://github.com/jsta/autodatadoc.

## Step 1: Create a csv data dictionary

Most people will use a spreadsheet program to manually create their data dictionary. An alternative approach is to use a more complex tool such as the [dataMeta](https://github.com/dmrodz/dataMeta) R package or [some other external tool](https://www.usgs.gov/products/data-and-tools/data-management/data-dictionaries) besides a spreadsheet program.

```{r, echo=FALSE}
dictionary <- dplyr::bind_rows(
  c(name = "country", description = "Country name"), 
  c(name = "year", description = "Year"), 
  c(name = "population", description = "Population") 
)
```

Our example data has a data dictionary that looks like this:

```{r, echo=FALSE}
knitr::kable(dictionary)
```

We should probably store our data dictionary under version control alongside our package. This is conventionally done by storing it in the `data-raw` folder.

## Step 2: Write a function to pass a tabular representation of our data dictionary as individual lines

We need to write a function to locate our data dictionary (in the `data-raw` folder), read the contents, pass it through our `tabular` function above, and return the output as individual roxygen lines:

```{r, eval=FALSE}
get_table_metadata <- function(path){
  dt <- read.csv(path, stringsAsFactors = FALSE)
  paste0(readLines(textConnection(tabular(dt))))
}

get_table_metadata("data-raw/dictionary.csv")
```

```{r, echo=FALSE}
get_table_metadata <- function(dt){
  paste0(readLines(textConnection(tabular(dt))))
}

get_table_metadata(dictionary)
```

## Ste 3: Automatically include this output in a roxygen block using the  `@eval` tag. 

To have this function evaluated during the course of roxygen building, we include the following in our code:

```r
#' @eval c("@details", get_table_metadata("data-raw/dictionary.csv"))
```

## Step 4: Run `roxygen::document` and/or `pkgdown::build_site`

Now when we build the documentation with roxygen::document() `@eval` will evaluate the function and include a nicely formatted data dictionary table in our docs (hooray!):

```{r, echo=FALSE}
library(printr)
?autodatadoc::population
```

## Further reading

https://www.usgs.gov/products/data-and-tools/data-management/data-dictionaries

https://github.com/r-lib/roxygen2/issues/645

https://blog.r-hub.io/2020/05/29/distribute-data/


